---
title: "HawkesFit tutorial: simulation, fitting and extension"
output: github_document
---
## Introduction

This package is designed for simulating and fitting the Hawkes processes and the HawkesN processes with several options of kernel functions. Currently, it assumes univariate processes without background event rates. Prior knowledge about the models is assumed in the following tutorial and please refer to [1] and [2] for details about the models.

```{r setup, include=FALSE}
library(HawkesFit)
```

## Simulating cascades

```{r}
lapply(seq(2), function(i) generate_Hawkes_event_series(c(K = 0.8, theta = 1), model_type = 'EXP'))
```

## Fitting a model on data


## Reference

[1] Rizoiu, Marian-Andrei, et al. "Hawkes processes for events in social media." Frontiers of Multimedia Research. Association for Computing Machinery and Morgan & Claypool, 2017.

[2] Rizoiu, Marian-Andrei, et al. "SIR-Hawkes: Linking epidemic models and Hawkes processes to model diffusions in finite populations." Proceedings of the 2018 World Wide Web Conference. International World Wide Web Conferences Steering Committee, 2018.
You can also embed plots, for example:
