---
title: "Profiling Twitter users with diffusions"
author: "Quyu Kong"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Profiling Twitter users with diffusions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(evently)
library(ggplot2)
library(tsne)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette gives a tutorial on profiling Twitter users with their initiated diffusions via modeling methods provided by `evently`. 
We use the dataset `auspol` shipped with `evently` in this task. It is a list of data frames (`r length(auspol)` in total) where each data frame represents a diffusion cascade. For each cascade, three fields are presented: `time` is the retweeting time relative to the original tweet (i.e., the first event); `user` is the corresponding Twitter user of the event;  `magnitude` is the number of followers the Twitter user has.

```{r}
head(auspol, n = 3)
```


```{r}
names(auspol) <- sapply(auspol, function(data) data$user[[1]])
auspol_group_fits_by_user <- group_fit_series(data = auspol, model_type = 'mPL', observation_times = Inf, cores = 10)

head(auspol_group_fits_by_user, n = 3)
```

```{r}
# keep only users with more than 5 cascades
auspol_group_fits_by_user <- Filter(function(x) length(x) >= 5, auspol_group_fits_by_user)
dist_matrix <- fits_dist_matrix(auspol_group_fits_by_user)

positions <- tsne(dist_matrix, k = 2)
```


```{r}
ggplot(as.data.frame(positions), aes(V1, V2)) +
  geom_point() +
  theme_void()
```

